Optimal (Multiway) Spatial Joins

### 2024-05-26 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$. We now explain how to find all the result tuples $\bm{t} \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$. Define $\J^\cat_i$ as the set of result tuples $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.

For every $\bm{t} = (r_1, r_2, ..., r_k) \in \J^\cat_k$, the top-left corner of $r_k$ must be covered by all of $r_1, ..., r_{k-1}$. This observation motivates us to find the target result tuples as follows. First, collect the set $P$ of top-left corners of all the rectangles in $R_k$. Remove from $P$ every point $p$ with the property that, there exists at least one $j \in [k-1]$ such that $p$ is covered by no rectangle in $R_j$. This requires solving $k-1$ instances of the detection version of Problem $\mathscr{A}$ (pairing $P$ together with a different $R_j$ where $j \in [k-1]$ as the input in each instance); the cost is $O(k n \log n)$ by the discussion in Section~\ref{sec:bricks}.

Let $P'$ be the set of remaining points in $P$ after the aforementioned removal. Next, for each $j \in [k-1]$, solve the reporting version of Problem $\mathscr{A}$ by feeding $P'$ and $R_j$ as the input. Doing so produces, for each point $p \in P'$, $\contained_{R_j}(p)$, which --- as defined in \eqref{eqn:contained} (noticing that $p$ is a degenerated ``horizontal segment'') --- is the set of rectangles in $R_j$ covering $p$. By the discussion in Section~\ref{sec:bricks}, the total cost of this step is bounded by
\myeqn{
    O\Big(kn \log n + \sum_{j \in [k-1]} \sum_{p \in P'} |\contained_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
}

We are now ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
\myeqn{
    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
}
